---
title: "Nationwide_exercise_AP"
author: "Annie_Phan"
date: "9/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Loading data to R

```{r}
par(mfrow=c(1,1))
RatesData<-read.csv(file=paste("Traindata_NW.csv",sep="/"),
           row.names=1,header=TRUE,sep=",")
#head(RatesData)
```

```{r}
# prediction y
Maturities<-RatesData[,80]
```

```{r}
# Visualization for predictor variables
RatesData_x<-data.matrix(RatesData[,1:78],rownames.force="automatic")
RatesData_x.ranks.matrix<-apply(RatesData_x,2,rank)
matplot(1:length(RatesData_x[,1]),RatesData_x,
        type="l",lty=1,
        xlab="",ylab="Rates")
```

```{r}
Covariance.Matrix<-cov(RatesData_x)
#Covariance.Matrix
```


```{r}
#cor(RatesData_x)
```

## 1. PCA assumptions:
The method is based on approximation of the covariance matrix. Thus, it has Gaussian assumption about the data.

Check if this assumption is correct.

```{r}
hist(RatesData_x[,3])
```

```{r}
qqnorm(RatesData_x[,3])
qqline(RatesData_x[,3])
```

The predictors in this dataset is Gaussian distribution.

## 2. Apply PCA

```{r}
PCA.RatesData<-princomp(RatesData_x)
names(PCA.RatesData)
```

```{r}
#summary(PCA.RatesData)
```

```{r}
plot(PCA.RatesData)
```

The plot clearly indicates that 5 components should be enough.
Make the loadings and factors.

```{r}
Loadings.5.factors<-PCA.RatesData$loadings[,1:5]
Factors.5.factors<-PCA.RatesData$scores[,1:5]
```

Create matrix of 5-factor PCA approximations.
Check approximation for one of the variables.

```{r}
Model.Matrix.5.factors<-t(apply(Factors.5.factors%*%t(Loadings.5.factors),1,"+",PCA.RatesData$center))
matplot(1:length(RatesData_x[,1]),cbind(RatesData_x[,6],Model.Matrix.5.factors[,6]),
        type="l",lty=1,col=c("black","red"),
        xlab=NULL,ylab=NULL)

```

Compare the covariance matrix of 78 variables with the model covariance matrix.

```{r}
#cov(Model.Matrix.5.factors)
```

```{r}
#Covariance.Matrix
```

Approximation is close.

Explore factors.

```{r}
matplot(1:length(Factors.5.factors[,1]),Factors.5.factors,
        type="l",lty=1,xlab=NULL,ylab="Factors",col=c("black","red"),
        main="First 5 factors",lwd=3)
```

```{r}
plot(Factors.5.factors,
     type="l",lty=1,xlab="Factor 1",ylab="Factor 2",
        main="First 2 factors",col=c("black","red"),lwd=3)
```

```{r}
cor(Factors.5.factors)
```

We see again long periods of persistent non-linear dependence.